---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r loading, echo=TRUE}
steps_dataset <- read.csv("./activity.csv", header = TRUE, na.strings = 'NA')
```

## What is mean total number of steps taken per day?

```{r date_summary, echo=TRUE}
date_summary <- aggregate(steps~date, steps_dataset, sum)
```
# Plot Histogram of steps taken each day:

```{r hist, echo=TRUE}
library(ggplot2)
ggplot(date_summary, aes(x=as.Date(date), y=steps))+
geom_col(fill='black')+
xlab('Date (Year = 2012)')+
ylab('Steps')+
ggtitle('Steps Taken per Day')
```

Calculate mean number of steps taken each day:

```{r mean, echo=TRUE}
mean(date_summary$steps)
```

The mean number of steps taken per day is 10,766.19 steps

Calculate median number of steps taken each day:

```{r median, echo=TRUE}
median(date_summary$steps)
```

The median number of steps taken per day is 10,765 steps

## What is the average daily activity pattern?



## Imputing missing values

Calculate number of NAs in dataset:

```{r na, echo=TRUE}
summary(steps_dataset$steps)
```

There are 2,304 NAs in the dataset

## Are there differences in activity patterns between weekdays and weekends?
